---
title: "Downloads"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

> Some additional details about the materials for this course that you can download and view them offline.

## Assessment 1 (Quiz)

<details><summary>**20214 (Oct 2021 - Feb 2022)**</summary>
<center>
<embed src="files/Assessment Repository/20214 (OCT 2021 - FEB 2022)/ASSESSMENT 1 (QUIZ)/QUESTIONS.pdf" type="application/pdf" width="100%" height="970px" />
<p>

If you can't view the pdf file, click on the file's link - <a href="files/Assessment Repository/20214 (OCT 2021 - FEB 2022)/ASSESSMENT 1 (QUIZ)/QUESTIONS.pdf"> Download PDF!</a>

</p>
</center>
</details>

<details><summary>**20204 (Oct 2020 - Feb 2021)**</summary>
<center>
<embed src="files/Assessment Repository/20204 (OCT 2020 - FEB 2021)/ASSESSMENT 1 (QUIZ)/QUESTIONS.pdf" type="application/pdf" width="100%" height="1000px" />
<p>

If you can't view the pdf file, click on the file's link - <a href="files/Assessment Repository/20204 (OCT 2020 - FEB 2021)/ASSESSMENT 1 (QUIZ)/QUESTIONS.pdf"> Download PDF!</a>

</p>
</center>
</details>

<details><summary>Answers</summary>
<center>
<embed src="files/Assessment Repository/20204 (OCT 2020 - FEB 2021)/ASSESSMENT 1 (QUIZ)/ANSWERS.pdf" type="application/pdf" width="100%" height="1000px" />
<p>

If you can't view the pdf file, click on the file's link - <a href="files/Assessment Repository/20204 (OCT 2020 - FEB 2021)/ASSESSMENT 1 (QUIZ)/ANSWERS.pdf"> Download PDF!</a>

</p>
</center>
</details>

## Assessment 2 (Test)

**20204 (Oct 2020 - Feb 2021)** [Assessment 2](files/Assessment%20Repository/20204%20(OCT%202020%20-%20FEB%202021)/ASSESSMENT%202%20(TEST)/QUESTIONS.pdf) [Answers](files/Assessment%20Repository/20204%20(OCT%202020%20-%20FEB%202021)/ASSESSMENT%202%20(TEST)/ANSWERS.pdf)

## Assessment 3 (Case Study - Project)

**20204 (Oct 2020 - Feb 2021)** [Assessment 3](files/Assessment%20Repository/20204%20(OCT%202020%20-%20FEB%202021)/ASSESSMENT%203%20(CASE%20STUDY-PROJECT)/CASE%20STUDY/QUESTIONS.pdf) [Answers](files/Assessment%20Repository/20204%20(OCT%202020%20-%20FEB%202021)/ASSESSMENT%203%20(CASE%20STUDY-PROJECT)/CASE%20STUDY/ANSWERS.pdf)

## Assessment 4 (Final Examination)  

<details><summary>**20214 (Oct 2021 - Feb 2022)**</summary>
```{r echo=FALSE, out.height='1000px', out.width='100%', message=FALSE, warning=FALSE}
knitr::include_graphics("files/Assessment Repository/20214 (OCT 2021 - FEB 2022)/FINAL/FINAL_20214.pdf")
```
</details>
If you can't view the pdf file, click on the file's link - [Download PDF!](files/Assessment Repository/20214 (OCT 2021 - FEB 2022)/FINAL/FINAL_20214.pdf)    

**20204 (Oct 2020 - Feb 2021)** [Assessment 4](files/Assessment%20Repository/20204%20(OCT%202020%20-%20FEB%202021)/ASSESSMENT%204%20(FINAL%20ASSESSMENT)/QUESTIONS.pdf)

## Past Final Examinations (Questions and Answers)  

<details><summary>**Dec 2018**</summary>
<center>
<embed src="files/Assessment Repository/Past Exams/2018 December.pdf" type="application/pdf" width="100%" height="1000px" />
</center>
</details>
If you can't view the pdf file, click on the file's link - [Download PDF!](files/Assessment Repository/Past Exams/2018 December.pdf)  

<details><summary>**June 2018**</summary>
<center>
<embed src="files/Assessment Repository/Past Exams/2018 June.pdf" type="application/pdf" width="100%" height="1000px" />
</center>
</details>
If you can't view the pdf file, click on the file's link - [Download PDF!](files/Assessment Repository/Past Exams/2018 June.pdf)  

<details><summary>**Jan 2018**</summary>
<center>
<embed src="files/Assessment Repository/Past Exams/2018 January.pdf" type="application/pdf" width="100%" height="1000px" />
</center>
</details>
If you can't view the pdf file, click on the file's link - [Download PDF!](files/Assessment Repository/Past Exams/2018 January.pdf)  

## Contributors {.appendix}

Kartini Kasim\
Senior Lecturer\
Department of Mathematical Sciences, UiTM Kedah